Bayesian Inference in Numerical Cognition: A Tutorial Using JASP

### 2019-12-09 jncTutorial/manuscript.Rmd
Some very small changes to linear regression

To compute the relevance of each covariate **after** data observation, we simply replace the prior model probabilities in the table by the posterior model probabilities. These are calculated using a generalisation of Equation~{\eqref{eqPriorToPosterior}} and specifies that the posterior model probability for model \( k=0, 1, \ldots p-1 = 3 \) is given by<!--
-->
\begin{align}
\label{eqPriorToPosteriorGeneral}
P(\mathcal{M}_{k} \mid d) = \frac{ \textnormal{BF}_{k0}(d) P(\mathcal{M}_{k})}{ \sum_{k=0}^{p-1} \textnormal{BF}_{k0}(d) P(\mathcal{M}_{k})},
\end{align}<!--
-->
where the comparison between the null model and itself is defined as \( \textnormal{BF}_{00}(d) = 1 \) regardless of the data. Suppose that the model \( \mathcal{M}_{k} \) has \( l \leq p \) active covariates, then for the Bayes factor \( \textnormal{BF}_{k0} \), we need to choose a prior on \( l + 2 \) parameters for \( \mathcal{M}_{k} \) and a prior on \( 2 \) parameters for \( \mathcal{M}_{0} \). @zellner1980posterior proposed to use essentially the same priors Jeffreys's used for his Bayesian \( t \)-test. This connection is made explicit by exploiting the assumption of normally distributed errors to rewrite the linear regression models as the following normal means models:<!--
-->
\begin{align}
\label{eqLinRegAsNormal0}
\mathcal{M}_{0} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0}, \sigma^{2}), \\
\mathcal{M}_{1} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0} + \beta_{1} x_{i1}, \sigma^{2}), \\
\mathcal{M}_{2} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0} + \beta_{2} x_{i2}, \sigma^{2}), \\
\label{eqLinRegAsNormal3}
\mathcal{M}_{3} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0} + \beta_{1} x_{i1} + \beta_{2} x_{i2} , \sigma^{2}), 
\end{align}<!--
-->
where \( Y_{i} \, | \, x_{i} \) refers to the outcome variable given the covariates \( x_{i} \).

More explicit, the Bayes factor \( \textnormal{BF}_{k0} \), where \( \mathcal{M}_{k} \) has \( l \) active covariates, uses a "uninformative" prior on the nuisance parameter \( \beta_{0} \) and \( \sigma \) in both models, and for the \( \mathcal{M}_{k} \) an additional \( l \)-dimensional multivariate Cauchy prior on the vector of population effect sizes \( \delta \). The vector of effect sizes is defined by rescaling the vector of active \( \beta \) coefficients with respect to \( \sigma \), but also with respect to the covariates. More explicitly, when \( l=1 \), say, only \( \beta_{2} \) is active then the effect size parameter is given by \( \delta = \beta_{2} s_{x_{2}}/\sigma \), where \( s_{x_{2}} \) is observed standard error of the second covariate. If the second covariate is standardised, then \( s_{x_{2}} = 1 \), and we get the same definition of effect size as in the \( t \)-test, but with \( \mu \) replaced by \( \beta_{2} \). Furthermore, with \( l = 1 \), the prior on \( \delta \) is than a \( 1 \)-dimensional Cauchy distribution, indeed, the same one we would use in the \( t \)-test.

>>>>

To compute the relevance of each covariate **after** data observation, we simply replace the prior model probabilities in the table by the posterior model probabilities. These are calculated using a generalisation of Equation~{\eqref{eqPriorToPosterior}} and specifies that the posterior model probability for model \( k=0, 1, \ldots p-1 = 3 \) is given by<!--
-->
\begin{align}
\label{eqPriorToPosteriorGeneral}
P(\mathcal{M}_{k} \mid d) = \frac{ \textnormal{BF}_{k0}(d) P(\mathcal{M}_{k})}{ \sum_{k=0}^{p-1} \textnormal{BF}_{k0}(d) P(\mathcal{M}_{k})},
\end{align}<!--
-->
where the comparison between the null model and itself is defined as \( \textnormal{BF}_{00}(d) = 1 \) regardless of the data. Suppose that the model \( \mathcal{M}_{k} \) has \( l \leq p \), say \( l =1 \), active covariates, then for the Bayes factor \( \textnormal{BF}_{k0} \), we need to choose a prior on \( l + 2 \) parameters for \( \mathcal{M}_{k} \) and a prior on \( 2 \) parameters for \( \mathcal{M}_{0} \). @zellner1980posterior proposed to use essentially the same priors Jeffreys's used for his Bayesian \( t \)-test. This can be made explicit by exploiting the assumption of normally distributed errors to rewrite the linear regression models as the following normal means models:<!--
-->
\begin{align}
\label{eqLinRegAsNormal0}
\mathcal{M}_{0} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0}, \sigma^{2}), \\
\mathcal{M}_{1} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0} + \beta_{1} x_{i1}, \sigma^{2}), \\
\mathcal{M}_{2} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0} + \beta_{2} x_{i2}, \sigma^{2}), \\
\label{eqLinRegAsNormal3}
\mathcal{M}_{3} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0} + \beta_{1} x_{i1} + \beta_{2} x_{i2} , \sigma^{2}), 
\end{align}<!--
-->
where \( Y_{i} \, | \, x_{i} \) refers to the outcome variable given the covariates \( x_{i} \).

More specifically, the Bayes factor \( \textnormal{BF}_{k0} \), where \( \mathcal{M}_{k} \) has \( l \) active covariates, uses a "uninformative" prior on the nuisance parameter \( \beta_{0} \) and \( \sigma \) in both models, and for the \( \mathcal{M}_{k} \) an additional \( l \)-dimensional multivariate Cauchy prior on the vector of population effect sizes \( \delta \). The vector of effect sizes is defined by rescaling the vector of active \( \beta \) coefficients with respect to \( \sigma \), but also with respect to the covariates. More explicitly, when \( l=1 \), say, only \( \beta_{2} \) is active then the effect size parameter is given by \( \delta = \beta_{2} s_{x_{2}}/\sigma \), where \( s_{x_{2}} \) is observed standard error of the second covariate. If the second covariate is standardised, then \( s_{x_{2}} = 1 \), and we get the same definition of effect size as in the \( t \)-test, but with \( \mu \) replaced by \( \beta_{2} \). Furthermore, with \( l = 1 \), the prior on \( \delta \) is than a \( 1 \)-dimensional Cauchy distribution, indeed, the same one we would use in the \( t \)-test.

### 2019-12-09 jncTutorial/manuscript.tex
Some very small changes to linear regression

To compute the relevance of each covariate \textbf{after} data observation, we simply replace the prior model probabilities in the table by the posterior model probabilities. These are calculated using a generalisation of Equation\textasciitilde{}\{\eqref{eqPriorToPosterior}\} and specifies that the posterior model probability for model \(k=0, 1, \ldots p-1 = 3\) is given by
\begin{align}
\label{eqPriorToPosteriorGeneral}
P(\mathcal{M}_{k} \mid d) = \frac{ \textnormal{BF}_{k0}(d) P(\mathcal{M}_{k})}{ \sum_{k=0}^{p-1} \textnormal{BF}_{k0}(d) P(\mathcal{M}_{k})},
\end{align}
where the comparison between the null model and itself is defined as \(\textnormal{BF}_{00}(d) = 1\) regardless of the data. Suppose that the model \(\mathcal{M}_{k}\) has \(l \leq p\) active covariates, then for the Bayes factor \(\textnormal{BF}_{k0}\), we need to choose a prior on \(l + 2\) parameters for \(\mathcal{M}_{k}\) and a prior on \(2\) parameters for \(\mathcal{M}_{0}\). Zellner and Siow (1980) proposed to use essentially the same priors Jeffreys's used for his Bayesian \(t\)-test. This connection is made explicit by exploiting the assumption of normally distributed errors to rewrite the linear regression models as the following normal means models:
\begin{align}
\label{eqLinRegAsNormal0}
\mathcal{M}_{0} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0}, \sigma^{2}), \\
\mathcal{M}_{1} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0} + \beta_{1} x_{i1}, \sigma^{2}), \\
\mathcal{M}_{2} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0} + \beta_{2} x_{i2}, \sigma^{2}), \\
\label{eqLinRegAsNormal3}
\mathcal{M}_{3} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0} + \beta_{1} x_{i1} + \beta_{2} x_{i2} , \sigma^{2}), 
\end{align}
where \(Y_{i} \, | \, x_{i}\) refers to the outcome variable given the covariates \(x_{i}\).

More explicit, the Bayes factor \(\textnormal{BF}_{k0}\), where \(\mathcal{M}_{k}\) has \(l\) active covariates, uses a \enquote{uninformative} prior on the nuisance parameter \(\beta_{0}\) and \(\sigma\) in both models, and for the \(\mathcal{M}_{k}\) an additional \(l\)-dimensional multivariate Cauchy prior on the vector of population effect sizes \(\delta\). The vector of effect sizes is defined by rescaling the vector of active \(\beta\) coefficients with respect to \(\sigma\), but also with respect to the covariates. More explicitly, when \(l=1\), say, only \(\beta_{2}\) is active then the effect size parameter is given by \(\delta = \beta_{2} s_{x_{2}}/\sigma\), where \(s_{x_{2}}\) is observed standard error of the second covariate. If the second covariate is standardised, then \(s_{x_{2}} = 1\), and we get the same definition of effect size as in the \(t\)-test, but with \(\mu\) replaced by \(\beta_{2}\). Furthermore, with \(l = 1\), the prior on \(\delta\) is than a \(1\)-dimensional Cauchy distribution, indeed, the same one we would use in the \(t\)-test.

>>>>

To compute the relevance of each covariate \textbf{after} data observation, we simply replace the prior model probabilities in the table by the posterior model probabilities. These are calculated using a generalisation of Equation\textasciitilde{}\{\eqref{eqPriorToPosterior}\} and specifies that the posterior model probability for model \(k=0, 1, \ldots p-1 = 3\) is given by
\begin{align}
\label{eqPriorToPosteriorGeneral}
P(\mathcal{M}_{k} \mid d) = \frac{ \textnormal{BF}_{k0}(d) P(\mathcal{M}_{k})}{ \sum_{k=0}^{p-1} \textnormal{BF}_{k0}(d) P(\mathcal{M}_{k})},
\end{align}
where the comparison between the null model and itself is defined as \(\textnormal{BF}_{00}(d) = 1\) regardless of the data. Suppose that the model \(\mathcal{M}_{k}\) has \(l \leq p\), say \(l =1\), active covariates, then for the Bayes factor \(\textnormal{BF}_{k0}\), we need to choose a prior on \(l + 2\) parameters for \(\mathcal{M}_{k}\) and a prior on \(2\) parameters for \(\mathcal{M}_{0}\). Zellner and Siow (1980) proposed to use essentially the same priors Jeffreys's used for his Bayesian \(t\)-test. This can be made explicit by exploiting the assumption of normally distributed errors to rewrite the linear regression models as the following normal means models:
\begin{align}
\label{eqLinRegAsNormal0}
\mathcal{M}_{0} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0}, \sigma^{2}), \\
\mathcal{M}_{1} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0} + \beta_{1} x_{i1}, \sigma^{2}), \\
\mathcal{M}_{2} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0} + \beta_{2} x_{i2}, \sigma^{2}), \\
\label{eqLinRegAsNormal3}
\mathcal{M}_{3} & : Y_{i} \, | \, x_{i} \sim \mathcal{N}(\beta_{0} + \beta_{1} x_{i1} + \beta_{2} x_{i2} , \sigma^{2}), 
\end{align}
where \(Y_{i} \, | \, x_{i}\) refers to the outcome variable given the covariates \(x_{i}\).

More specifically, the Bayes factor \(\textnormal{BF}_{k0}\), where \(\mathcal{M}_{k}\) has \(l\) active covariates, uses a \enquote{uninformative} prior on the nuisance parameter \(\beta_{0}\) and \(\sigma\) in both models, and for the \(\mathcal{M}_{k}\) an additional \(l\)-dimensional multivariate Cauchy prior on the vector of population effect sizes \(\delta\). The vector of effect sizes is defined by rescaling the vector of active \(\beta\) coefficients with respect to \(\sigma\), but also with respect to the covariates. More explicitly, when \(l=1\), say, only \(\beta_{2}\) is active then the effect size parameter is given by \(\delta = \beta_{2} s_{x_{2}}/\sigma\), where \(s_{x_{2}}\) is observed standard error of the second covariate. If the second covariate is standardised, then \(s_{x_{2}} = 1\), and we get the same definition of effect size as in the \(t\)-test, but with \(\mu\) replaced by \(\beta_{2}\). Furthermore, with \(l = 1\), the prior on \(\delta\) is than a \(1\)-dimensional Cauchy distribution, indeed, the same one we would use in the \(t\)-test.
